Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-11-13 paper/Interpreting Random Forest Predictions for Firearm Identification Using LIME.Rmd
changed paper file name

# Load libraries
library(tidyverse)
library(bulletr)
library(randomForest)
library(lime)
library(cowplot)
```

@@

  # Apply the lime function from the lime package (with a seed set)
  # Note that the as_classifier must be added since rtrees is from the
  # randomForest package and not fit using caret or one of the other 
  # available models specified in the lime package. Additionally, the
  # randomForest package must be loaded in order to run the functions
  # from the lime package. (I should check on exactly how this works.)
  set.seed(84902)
  hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                        model = as_classifier(rtrees),
                        n_bins = 6)

@@

```{r}
# Function for writing bins given the bin boundaries
write_bins <- function(bin_data){

  # Compute the number of bins based the input dataframe
  nbins = length(bin_data) - 2

  # Create an empty matrix to store the bins in
  row <- matrix(NA, nrow = 1, ncol = nbins + 1)

  # Put the feature name in the first column
  row[,1] <- as.character(bin_data[[1]])

  # Put the 1st through 2nd to last bins in the 2nd through 2nd to last columns
  row[,2:nbins] <- sapply(2:nbins, 
                          FUN = function(number) sprintf("[%.2f, %.2f)", 
                                                         as.numeric(bin_data[number]),
                                                         as.numeric(bin_data[number + 1])))

  # Fill in the last column with the upper bin (based on whether it should be infinite or not)
  row[,nbins + 1] <- ifelse(is.na(bin_data[nbins + 2]),
                            sprintf("(%.2f, %s)", as.numeric(bin_data[nbins + 1]), "\U221E"),
                            sprintf("(%.2f, %.2f]", as.numeric(bin_data[nbins + 1]),
                                    as.numeric(bin_data[nbins + 2])))

  return(row)

@@

  # Assign appropriate names to the bin columns
  names(hamby_bins) <- c("Feature", "Lower Bin", 
                         sapply(1:(nbins-2), function(bin) sprintf("Middle Bin %.0f", bin)),
                         "Upper Bin")

@@

```{r}
# Function to use for creating bin labels in the test_explain dataset
bin_labeller <- function(feature, feature_value, bin_data = hamby_divisions){

  if (is.na(feature)) {

    # Set feature_bin to NA if feature is NA
    feature_bin <- NA

    # Subset the bin cuts table to the selected feature
    feature_bin_data <- bin_data %>%
      filter(Feature == feature) %>%
      select(-Feature, -Lower, -Upper)

    # Compute the number of bins based the input dataframe
    nbins = length(bin_data) - 2

    # Determine which bin the case falls in
    if(feature_value <= feature_bin_data[1]){
      feature_bin <- paste(feature, "(lower bin)")
    } else if (feature_bin_data[nbins - 1] < feature_value){
      feature_bin <- paste(feature, "(upper bin)")
    } else {
      middle_bin_checks <- data.frame(middle_bin_number = 1:(nbins-2),
                                      contained = sapply(1:(nbins - 2),
                                                         FUN = function(number) {
                                                           feature_bin_data[number] < feature_value &
                                                           feature_value <= feature_bin_data[number + 1]}))
      feature_bin <- sprintf("%s (middle bin %.0f)", 
                             feature,
                             middle_bin_checks %>% filter(contained == TRUE) %>% pull(middle_bin_number))
    }

  }

  # Return the bin
  return(feature_bin)

@@

  # Join the data and the explanations and edit and add additional variables
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature = factor(feature),
           feature_bin = mapply(bin_labeller, feature = feature, 
                                feature_value = feature_value)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE)

@@

```{r}
check <- hamby224_test_explain %>%
  select(-data, -prediction) %>%
  group_by(case) %>%
  slice(1) %>%
  select(case, set, rfscore, model_prediction, model_r2) %>%
  mutate(diff = rfscore - model_prediction,
         mean = (rfscore + model_prediction) / 2)

sum((subset(check$diff, check$set == "Set 1"))^2, na.rm = TRUE)
sum((subset(check$diff, check$set == "Set 11"))^2, na.rm = TRUE)
```

@@

>>>>

# Load libraries
library(tidyverse)
library(bulletr)
library(randomForest)
library(lime)
library(cowplot)

# Source functions
source("../code/helper_functions.R")
```

@@

  # Apply the lime function from the lime package (with a seed set)
  # Note that the as_classifier must be added since rtrees is from the
  # randomForest package and not fit using caret or one of the other 
  # available models specified in the lime package. Additionally, the
  # randomForest package must be loaded in order to run the functions
  # from the lime package. (I should check on exactly how this works.)
  set.seed(84902)
  hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                        model = as_classifier(rtrees),
                        n_bins = 4)

@@

@@

  # Assign appropriate names to the bin columns
  if (nbins == 2){
    names(hamby_bins) <- c("Feature", "Lower Bin", "Upper Bin")
  } else {
    names(hamby_bins) <- c("Feature", "Lower Bin", 
                           sapply(1:(nbins-2), 
                                  function(bin) sprintf("Middle Bin %.0f", bin)),
                           "Upper Bin")
  }

@@

@@

  # Join the data and the explanations and edit and add additional variables
  # Create the feature bin labels using my function "bin_labeller"
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature_bin = mapply(bin_labeller, feature = feature, 
                                feature_value = feature_value)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE, 
           feature = factor(feature))

@@

```{r}
check <- hamby224_test_explain %>%
  select(-data, -prediction) %>%
  group_by(case) %>%
  slice(1) %>%
  ungroup() %>%
  select(case, set, rf_features, rfscore, model_prediction, model_r2) %>%
  mutate(diff = rfscore - model_prediction,
         mean = (rfscore + model_prediction) / 2)

sum((subset(check$diff, check$set == "Set 1"))^2, na.rm = TRUE)
sum((subset(check$diff, check$set == "Set 11"))^2, na.rm = TRUE)

check %>%
  filter(set == "Set 1") %>%
  select(rf_features, diff, -case) %>%
  gather(key = feature, value = feature_value, rf_features) %>%
  select(feature, feature_value, diff) %>%
  ggplot(aes(x = feature_value, y = diff)) + 
  geom_point() + 
  facet_wrap( ~ feature, scales = "free") + 
  geom_hline(yintercept = 0, color = "blue")
```

@@

Ideas for improvement of LIME: 
  - change the binning method
